---
title: "p8105_hw5_jl5548"
author: "J L"
date: "November 7, 2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem 1

The code chunk below loads the iris dataset from the tidyverse package and introduces some missing values in each column. The purpose of this problem is to fill in those missing values.

```{r}
library(tidyverse)

set.seed(10)

iris_with_missing = iris %>% 
  map_df(~replace(.x, sample(1:150, 20), NA)) %>%
  mutate(Species = as.character(Species))
```

There are two cases to address:

* For numeric variables, you should fill in missing values with the mean of non-missing values
* For character variables, you should fill in missing values with "virginica"

Write a function that takes a vector as an argument; replaces missing values using the rules defined above; and returns the resulting vector. Apply this function to the columns of iris_with_missing using a map statement.

```{r}
## write a function that takes a vector as an argument, replace missing values using rules defined above and returns the resulting vector
fill = function(x){
  if(is.numeric(x)){
    x = x %>% 
      replace_na(mean(x[!is.na(x)])) %>% 
      round(digits = 1)
  } else if (is.character(x)){
    x = x %>% 
      replace_na("virginica")
  }
  x
}

## apply this function to columns of the dataset
iris_with_missing = map_df(.x = iris_with_missing, ~ fill(.x))

## show the dataset
iris_with_missing
```



## Problem 2

Create a tidy dataframe containing data from all participants, including the subject ID, arm, and observations over time:

* Start with a dataframe containing all file names; the list.files function will help

```{r}
df_name = list.files(path = "./data") %>%  # create a vector containing all file names
  as.data.frame() %>%   # turn the vector into a dataframe
  rename("filename" = ".")  # rename the column as "filename"

# display the dataframe
df_name
```

* Iterate over file names and read in data for each subject using purrr::map and saving the result as a new variable in the dataframe

```{r}
df_long = df_name %>% 
  # use map to read data into a nested column "data"
  mutate(data = map(.x = paste0("./data/", pull(df_name, filename)), read_csv)) %>% 
  # unnest the column "data"
  unnest()

# display the dataframe
df_long
```

* Tidy the result; manipulate file names to include control arm and subject ID, make sure weekly observations are “tidy”, and do any other tidying that’s necessary

```{r}
df_long_tidy = df_long %>% 
  # separate the the column filename into two columns:  "arm" and "subject_id"
  separate(filename, into = c("arm", "subject_id"), sep = "_") %>% 
  mutate(
    # remove ".csv" at the end of each subject_id
    subject_id = str_remove(subject_id, pattern = ".csv"),
    # replace "con" as "control" 
    arm = str_replace(arm, "con", "control"),
    # replace "exp" as "experimental"
    arm = str_replace(arm, "exp", "experimental")
  )

# display the dataframe
df_long_tidy
```

