---
title: "p8105_hw5_jl5548"
author: "J L"
date: "November 7, 2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem 1

The code chunk below loads the iris dataset from the tidyverse package and introduces some missing values in each column. The purpose of this problem is to fill in those missing values.

```{r}
library(tidyverse)

set.seed(10)

iris_with_missing = iris %>% 
  map_df(~replace(.x, sample(1:150, 20), NA)) %>%
  mutate(Species = as.character(Species))
```

There are two cases to address:

* For numeric variables, you should fill in missing values with the mean of non-missing values
* For character variables, you should fill in missing values with "virginica"

Write a function that takes a vector as an argument; replaces missing values using the rules defined above; and returns the resulting vector. Apply this function to the columns of iris_with_missing using a map statement.

```{r}
## write a function that takes a vector as an argument, replace missing values using rules defined above and returns the resulting vector
fill = function(x){
  if(is.numeric(x)){
    x = x %>% 
      replace_na(mean(x[!is.na(x)])) %>% 
      round(digits = 1)
  } else if (is.character(x)){
    x = x %>% 
      replace_na("virginica")
  }
  x
}

## apply this function to columns of the dataset
iris_with_missing = map_df(.x = iris_with_missing, ~ fill(.x))

## show the dataset
iris_with_missing
```



## Problem 2

Create a tidy dataframe containing data from all participants, including the subject ID, arm, and observations over time:

* Start with a dataframe containing all file names; the list.files function will help

```{r}
df_name = list.files(path = "./data") %>%  # create a vector containing all file names
  as.data.frame() %>%   # turn the vector into a dataframe
  rename("filename" = ".")  # rename the column as "filename"

# display the dataframe
df_name
```

* Iterate over file names and read in data for each subject using purrr::map and saving the result as a new variable in the dataframe

```{r}
df_long = df_name %>% 
  # use map to read data into a nested column "data"
  mutate(data = map(.x = paste0("./data/", pull(df_name, filename)), read_csv)) %>% 
  # unnest the column "data"
  unnest()

# display the dataframe
df_long
```

* Tidy the result; manipulate file names to include control arm and subject ID, make sure weekly observations are “tidy”, and do any other tidying that’s necessary

```{r}
df_long_tidy = df_long %>% 
  # separate the the column filename into two columns:  "arm" and "subject_id"
  separate(filename, into = c("arm", "subject_id"), sep = "_") %>% 
  mutate(
    # remove ".csv" at the end of each subject_id
    subject_id = str_remove(subject_id, pattern = ".csv"),
    # replace "con" as "control" 
    arm = str_replace(arm, "con", "control"),
    # replace "exp" as "experimental"
    arm = str_replace(arm, "exp", "experimental")
  )

# display the dataframe
df_long_tidy
```

Make a spaghetti plot showing observations on each subject over time, and comment on differences between groups.

```{r}
# create a dataframe for the plot
df_plot = df_long_tidy %>% 
  # convert all observations from week 1 to week 8 into one column "observation" with the label column "week"
  pivot_longer(
    cols = week_1:week_8,
    names_to = "week",
    values_to = "observation"
  )
# make the plot
plot_long = df_plot %>% 
  # set values for x and y axes, and different colors according to subject_id
  ggplot(aes(x = week, y = observation, color = subject_id)) +
  # add lines to connect observations through 8 weeks for each subject
  geom_line(aes(group = subject_id)) + 
  # separate by control and experimental arms
  facet_grid(.~arm) + 
  # add titles
  labs(
    title = "observations on subjects over 8 weeks in control and experimental arms",
    x = "week",
    y = "observation"
  ) +
  # adjust the x_axis texts to make it readable
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

# display the plot
plot_long
```

As shown in the plot, at the first week, the experimental arm has overall higher observation value than the control arm. Furthermore, the experimental arm experienced an overall increase in observation value which did not show in the control arm. 



## Problem 3

In this problem, you will conduct a simulation to explore power in a simple linear regression.

First set the following design elements:

* Fix n=30
* Fix xi1 as draws from a standard Normal distribution
* Fix β0=2
* Fix σ2=50

Set β1=0. Generate 10000 datasets from the model yi = β0 + β1xi1 + ϵi with ϵi∼N[0,σ2]. For each dataset, save β̂1 and the p-value arising from a test of H:β1=0 using α=0.05. Hint: to obtain the estimate and p-value, use broom::tidy to clean the output of lm.

```{r}

```

